---
title: "Biodiversity Module"
author: <a href="http://bigdata.duke.edu/data">Duke Data+</a> Interactive Environmental Data Team
date: ""
output: 
  html_document:
    theme: journal
runtime: shiny
---
```{r, echo=FALSE}
suppressPackageStartupMessages(library(shiny))
suppressPackageStartupMessages(library(leaflet))
suppressPackageStartupMessages(library(rgbif))
```


```{r, echo=FALSE}
mythical<- read.csv("data/mythical.csv")
m<- as.character(mythical[,1])
getCommunity <- function(S=10,N=1000,equalprob=T){
  if(S > 51) S <- 51 # right now we only have 26 species - each unique letter in the alphabet
  prob <- rep(1/S,S)
  if(!equalprob){
    Nx <- rpois(S,sample(1:S,1))
    prob <- Nx/sum(Nx)    
  }
  sub<-m[sample(S,N,replace=T,prob=prob)]
  print(sub)
  return (sub)
}
```

```{r,echo=FALSE}
ui<- fluidPage(
  fluidRow(
    h4("Click 'Get Community' to view the community distribution. Click 'Add To Table' to store values in a table."),
    br(),
    
    #slider input for Tab 2 
    column(3, sliderInput('num_species', 'Number of Species to Examine: ', min=2, max=51, value=12),
           
    #numeric input for number of individuals
        numericInput('num_individuals', 'Number of individuals in the Population: ', min=1000, max=10000, value=1000, step=50),
    #buttons controlling "skewed" or "uniform" and "Get Community" and "Save to Table"
        radioButtons("random", label= "Type of community distribution",choices=c("uniform"="yes", "skewed"="no"), selected="yes"),
        actionButton("get", label = "Get Community"),
        actionButton("add", label= "Save to Table")),
    
#creates space for multiple choice questions
    column(5, radioButtons(
                 "Q1", 'What happens to the richness if you have a very high number of species oberservations?',
                 choices = c("It Decreases"=1, "It stays the same"=2,"It Increases"=3), selected=F
               ),
               h5(textOutput("ansQ1"), style="color:blue"),
#creates space for the distribution plot          
            plotOutput("dist"),
        tableOutput("dataLine")
  ),
#question 2 multiple choice 
    column(4, radioButtons(
      "Q2", 'What happens to the Simpson Index if you keep the number of species at 20 and increase the population size from very small to very large?',
                 choices = c("It Decreases"=1, "It stays the same"=2,"It Increases"=3), selected=F
               ),
               h5(textOutput("ansQ2"), style="color:blue"),
      #creates space for table output
           tableOutput("dataTable"))
    )
    
  )
        

server <-  function(input,output){
# output for Q1     
  output$ansQ1<- reactive({
    validate( need(input$Q1, "Select an answer") )
if (input$Q1==3){answer<- "Correct! Richness is proportional to the number of species."}
    else{answer<-"Incorrect"}
    return(answer)
    })
 #Output for Q2 multiple choice 
    output$ansQ2<- reactive({
    validate( need(input$Q2, "Select an answer") )
if (input$Q2==1){answer<- "Correct! Why do you think this happens?"}
    else{answer<-"Incorrect"}
    return(answer)
    })
    
      S<-reactive({input$num_species})
      N<-reactive({input$num_individuals})
      ran<- reactive({
        if (input$random == "yes") {TRUE}
        else {FALSE}
        })
 
      values <- reactiveValues(df=matrix(NA,nrow=0,ncol=4))
#create barplot based on values chosen by slider input/entry on Tab2     
    species <- eventReactive(input$get,{getCommunity(S(),N(),ran())})
    output$dist= renderPlot({
      par(mar=c(12,4,2,0.4))
      barplot(table(species()), main="Species observations",xlab="", ylab="Number of Individuals",las=2)
    })
    
#calculate data metrics based on numbers chosen by slider input/entry on Tab2    
    metrics<- eventReactive(input$get,{  
      richness<- length(unique(species()))
      Pi <- table(species())/length(species())
      shannon<-round((-sum(Pi*log(Pi))),2)
      print (shannon)
      lambda <- round(sum(Pi^2),2)
      simpson<-lambda
      num_s<- input$num_species
      num_i<- input$num_individuals
      x<- c(num_i,richness, shannon, simpson)
      x
      })
    
#Add generated data from app to the table
  adddata <- observeEvent(input$add,{       
     newLine <- c(metrics()[1],metrics()[2],metrics()[3],metrics()[4])
     values$df <- rbind(values$df, newLine)})
  
  output$dataLine <- renderTable({
    tmp <- matrix(unlist(metrics()))
    rownames(tmp) <- c("N","Richness (S)","Shannon","Simpson")
    colnames(tmp) <- NULL
    tmp
  })
#Output the data table comprised of data generated from the app
  output$dataTable <- renderTable({
    tmp<- values$df
    colnames(tmp)<- c("N", "S", "Shannon", "Simpson")
    rownames(tmp)<-  NULL
    tmp})
      } 

```

```{r, echo=FALSE}
#ui for the Real Life Analysis
ui2<-  (
  fluidPage(
    fluidRow(
      h3("To Begin: Zoom in to any location and click 'Update Data'.
      Scroll down to see the real biodiversity metrics for the given view"),
 column(5, br(),
        selectInput("type","Select what GBIF data to retrieve:",
              choices = c(
              "Birds"=212,
              "Mammals"=359,
              "Reptiles"= 358,
              "Plants"=6            
                )), 
        br(),
        selectInput("obs_method","Method Of Observation",
              choices= c(
                "Human Observation"="HUMAN_OBSERVATION",
                "Fossil Specimen"="FOSSIL_SPECIMEN",
                "Machine Observation"="MACHINE_OBSERVATION")
                )),
  column(5, offset= 1, 
         br(),
         sliderInput("limit", "Retrieve how many results?", min=500, max=5000, value=50), 
         br(),
         actionButton("update", "Update Data")),
         br(),
  
 leafletOutput("map"), br(),br(),
 textOutput("error"),

    #mainPanel(
      tabsetPanel(type = "tabs", 
        tabPanel("Data Table", tableOutput("op")),
        tabPanel("Rank Abundance Curve", plotOutput("raCurve")),
        tabPanel("Biodiversity Metrics", textOutput("richness"), textOutput("shannon"), textOutput("simpson"), textOutput("invS"),textOutput("indv"), textOutput("spec"))
      )
  
  )))

#server for Real Life Analysis
server2<- function(input,output){
  #retrieving proper map tiles
   url<- reactive({paste( "http://api.gbif.org/v1/map/density/tile?x={x}&y={y}&z={z}&type=TAXON&key=",input$type,"&palette=reds&resolution=4", sep="")})
  #Rendering biodiversity map
  output$map= renderLeaflet({leaflet() %>% addTiles() %>% setView(lng=-79,lat=36,zoom=3) %>%
  addTiles(urlTemplate=url())
 })
  
#retrieving coordinate window visible in leaflet
datInBounds <- reactive({

    bounds <- input$map_bounds
    latRng <- sort(round(range(bounds$north, bounds$south),5))
    lngRng <- sort(round(range(bounds$east, bounds$west),5))
    if(latRng[1] < -90) latRng[1] <- -90
    if(latRng[2] > 90) latRng[2] <- 90
    if(lngRng[1] < -180) lngRng[1] <- -180
    if(lngRng[2] > 180) lngRng[2] <- 180
    
    lats <- paste(latRng[1],latRng[2],sep=",")
    lons <- paste(lngRng[1],lngRng[2],sep=",")
    coords<- c(lats,lons)
    if (is.null(input$map_bounds)){
      lats <- "34,36"
      lons <- "-80,-77"
    }
    return (coords)})  
  
#taking the coordinates and getting the observations for that window
gbf <- eventReactive(input$update,{
  if (input$map_zoom < 8) {
    output$error= renderText({"error, please zoom in"})
    return (NULL)}
  else{
    output$error=renderText({("")})
    coords<- datInBounds() 
    occ_search(taxonKey=input$type, limit=input$limit, hasCoordinate=TRUE, spatialIssues=FALSE, basisOfRecord=input$obs_method,decimalLatitude=coords[1], decimalLongitude=coords[2], fields=c('species','basisOfRecord','decimalLatitude', 'decimalLongitude'), return='data')}
    
    })

#calculating the biodiversity metrics
metrics<- eventReactive(input$update, {
    data<- gbf()
    richness<- length(unique(data$species))
    Pi <- table(gbf()$species)/length(data$species)
    shannon<-round((-sum(Pi*log(Pi))),2)
    lambda <- round(sum(Pi^2),2)
    simpson<-lambda
    invS<- 1/lambda
    m<- c(richness, shannon, simpson, invS)
    m
    # output$richness= renderPrint({m[1]})
    #output$shannon= renderPrint({shannon})
    #output$simpson= renderPrint({simpson})
})

#rendering plot and table outpets
output$op= renderTable({data.frame(species=sort(table(gbf()$species),decreasing=T))}) 
output$raCurve= renderPlot({plot(sort(table(gbf()$species),decreasing=T),xlab="Rank",ylab="Abundance")})

#render the live biodiverity metrics
output$richness= renderPrint(paste("Richness:", {metrics()[1]}))
output$shannon= renderPrint(paste("Shannon Index:", {metrics()[2]}))
output$simpson= renderPrint(paste("Simpson Index:", {metrics()[3]}))
output$invS= renderPrint(paste("Inverse Simpson Index: ", {metrics()[4]}))
output$indv = renderPrint(paste("Number of Individuals:", {nrow(gbf())}))
#also dim(gbf[1])
output$spec = renderPrint(paste("Number of Species:", {length(unique(gbf()$species))}))

}


```

```{r,echo=FALSE}
tabsetPanel(
  tabPanel("About"),
  tabPanel("Experimental Analysis", shinyApp(ui=ui, server=server, options=list(height=1000))),
  tabPanel("Real Life Analysis", shinyApp(ui=ui2, server=server2, options = list(height=1000)))
)

```

## Background

Biodiversity indices are used by environmental scientists to measure the diversity of species in a community.

Three common measures are:

* Richness
* Shannon index
* Simpson index

We can calculate these indices with a data set of observations. For example, here is a hypothetical community (for new communities, just reload this page):

```{r, echo=FALSE}
dat <- sample(m,20,replace=T)
dat
```

## Richness

Richness ($S$) is the number of unique species. With a list of organism observations (above), the richness is:

```{r, echo=FALSE}
S <- length(unique(dat))
paste(S, "unique species")
```


## Shannon index

The Shannon index measures the evenness of species observations with relative proportions. The index is:

$H'=- \sum_{i=1}^S p_i \ln p_i$

where $S$ is the number of unique species and $p_i$ is the proportion of the $i^{th}$ species, which is $\frac{n_i}{N}$, where $N$ is the total number of individuals.


```{r, echo=FALSE}
Pi <- table(dat)/length(dat)

H <- -sum(Pi*log(Pi))
H
```

Higher values have higher evenness, and the maximum value is the log of the species richness ($\ln S$).   For this example: the maximum value is:

```{r, echo=FALSE}
X<- log(S)
X
```

Therefore, this population has a high biodiversity

## Simpson index

The Simpson index measures the concentration of observations in certain groups. This index is:

$\lambda=\sum_{i=1}^S p_i^2$

It measures the probability that two individuals randomly chosen from a sample will have the same species.

For our example data set:

```{r, echo=FALSE}
lambda <- sum(Pi^2)
paste("Simpson is:", lambda)
```

It ranges between 0 and 1, and lower numbers are more diverse. Because of this inverse ranking, some people refer to the inverse of the Simpson index:

$D = 1/\lambda$

```{r, echo=FALSE}
D <- 1/lambda
paste("Inverse Simpson is:",  D)
```

Here, 1 is the lowest number and higher values have higher diversity.


